---
title: "Linear Regression Age Vs. BMI"
author: "Haruma Kurita"
date: "2025-10-28"
output: html_document
---

## Introduction

This project investigates whether age affects Body Mass Index (BMI) using the NHANES (National Health and Nutrition Examination Survey) dataset.

### Research Question
Does age have a significant relationship with BMI?
```{r, results='hide'}
# Load the libraries
library(ggplot2)
library(NHANES)
library(tidyverse)
```

## Data Exploration
```{r}
# Inspect the data
head(NHANES)
nrow(NHANES)

# Check the NA
sum(is.na(NHANES$Age))
sum(is.na(NHANES$BMI))
```

## Data Cleaning
```{r}
# Removing the NA and creating a new data frame
data_clean <- NHANES |> 
  filter(!is.na(Age) & !is.na(BMI)) |> 
  select(Age, BMI)
sum(is.na(data_clean))
```

## Linear Regression Analysis
```{r}
# Making a linear model
model <- lm(BMI ~ Age, data = data_clean)
summary(model)

# Making a scatterplot
ggplot(data_clean, aes(x = Age, y = BMI)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Relationship Between Age and BMI",
    x = "Age in Years",
    y = "BMI"
  ) +
  theme_minimal()
```

The R^2^ value from the linear model was 0.1664, a p-value less than 0.001, and the F-statistic was 1922. This indicates that that age does have an effect on BMI with highly statistically significant values from the tests. Age has a moderate effect on BMI but the other 83% of the BMI variation comes from other factors.


